### 2021-04-12 paper/parlr_paper_1.tex
First draft of Efficient Computation

\setstretch{1}
\begin{flushright} 
    \end{flushright}
    \begin{center} \textbf{Base Model and New Linkage Cluster Model Specification}

@@

@@

(This will be revised to talk more expliticly about sampling \(Z_j\)
independently)

@@

\[\lambda \sim \text{Beta}(\alpha_{\lambda}, \beta_{\lambda}) \] The
prior for \(Z_j\) has equal probability of matching to all records
\(i\in A\), and non-matching probability governed by \(\lambda\).
Therefore a \(\lambda \sim \text{Beta}(1, 1)\) corresponds to a prior
belief that nonmatches and matches are equally likely, and a
\(\lambda \sim \text{Beta}(1, \frac{1}{n_A})\) prior corresponds to a
uniform prior on the labelling of \(\mathbf{Z}\).

@@

\underline{Sample $\lambda^{(s+1)}|\mathbf{Z}^{(s)}$:} As a function of
\(\lambda\), the linkage structure parameter \(\mathbf{Z}\) is sequence
of successes (when \(z_j < n_A + 1\)) and failures (when
\(z_j = n_A + 1\)), and therefore
\(p(\mathbf{Z}|\lambda) = \mathcal{L}(\lambda|\mathbf{Z})\) is
determined only by the number of dupliates
\(D = \sum_{i=1}^{n_B}\mathbf{1}_{z_j < n_A + 1}\) encoded by
\(\mathbf{Z}\). Thus we have

@@

\underline{Sample $\mathbf{Z}^{(s+1)}|\Gamma, \mathbf{m}^{(s+1)}, \mathbf{u}^{(s+1)}, \lambda^{(s+1)}$:}
Because we sample \(Z_j\) independently of all other \(Z_{j'}\), we use
only the full conditional for an individual \(Z_j\). Let \(\Gamma_{.j}\)
denote the set of \(n_A\) comparison vectors with \(j \in B\), and note
that as a function of \(Z_j\), the likelihood
\(p(\Gamma_{.j}|Z_j, \mathbf{m}, \mathbf{u}) = \mathcal{L}(Z_j|\Gamma_{.j}, \mathbf{m}, \mathbf{u})\)
is a discrete distribution with probalities proportional to

@@

In order to make fair comparisons against the Sadinle 2017 model, we
integrate over the posterior of \(\lambda\) and rearrange terms to
produce the final full conditoinal:

\[Z_j^{(s+1)} | \mathbf{m}, \mathbf{u}, \mathbf{Z^{(s)}} \propto
\begin{cases} 
    w_{ij}  & z_j \leq n_A; \\
     n_A \frac{n_B - D + \beta_{\lambda}}{D + \alpha_{\lambda}} &  z_j  = n_A + 1 \\
\end{cases}\]

\hypertarget{decouplerecouple}{%
\subsection{Decouple/Recouple}\label{decouplerecouple}}

(This will be revised; I'm not sure if I really believe the ``Recouple''
step anymore! I'll need look into it more on larger datasets)

One difficulty in record linkage is conducting linkage while taking
appropriate dependencies into account. In the context of this paper in
which we assume that each file is recorded without duplicates, this
means our linkage procedure should respect one-to-one matching. The
posterior sampling of \(\mathbf{Z}\) outlined above resolves half of
this issue; since we only sample one potential match for record \(j\) in
each Gibbs iteration, it is impossible to match \(j\) to two records at
once. This in itself is an improvement over the classic Fellegi-Sunter
method that labels record pairs as matches or non matches independently
of all other pairs. However, our posterior sampling of \(\mathbf{Z}\)
does allow for the possibility that \(Z_j = i = Z_{j'}\) for some
\(j' \neq j\), which cleary violates the assumption that
\(\mathbf{X}_B\) is recorded without duplicates. Sadinle 2017 resolves
this issue by adapting the set of candidate matches for every \(Z_j\),
but we forgo this rigidity and instead ``decouple'' the linkage vector
\(\mathbf{Z}\) for the computational advantantages of sampling \(Z_j\)
independently. In our formulation, such illegal matches occasionally
occur within individual Gibbs iterations due to simple randomness, but
they do not occur consistently enough between the same records to
measurably influence the eventual Bayes estimate of the linkage
structure.

However, if we want to ensure the theoretic purity of each MCMC sample
of the linkage structure, we can ``recouple'' \(\mathbf{Z}\) by simply
resolving these illegal matchings. Explictly, define sets
\(R(i) := \{j | Z_j = i\}\), and define all sets such that
\(|R(i)| > 1\) as \emph{resolution sets}. Within each set, resample the
candidate matches according to their Fellegi-Sunter weights. We show
this step to be computationally inexpensive and to be a reasonable
replication of more organic one-to-one matching. This procedure would be
most useful in cases where there are many ``near matches'' across files,
but in most cases, we find this this step to be unnecessary.

@@

Same as Sadinle.

\hypertarget{efficient-computation}{%
\subsection{Efficient Computation}\label{efficient-computation}}

Adopting some from Enamorado 2019, some steps my own. very promising,
stay tuned!

\newpage

\hypertarget{proposed-linkage-cluster-model}{%
\subsection{Proposed Linkage Cluster
Model}\label{proposed-linkage-cluster-model}}

(Still thinking through best notation and terminology)

The primary contribution of this paper is the establishment of linkage
clusters, which allow the reliability parameters \(\mathbf{m}^f\), the
discernment parameters \(\mathbf{u}^f\), and also the linkage
probability parameter \(\lambda\) to vary according to certain
attributes in the record pairs. We shall see that this set up upweights
linkage within clusters and downweights linkage across clusters, thus
using attribute information in a way not yet explored in the literture.

Partition files \(A\) and \(B\) into \emph{record clusters}
\(\{A_1, \ldots, A_k\}\) and \(\{B_1, \ldots, B_k\}\) according to some
attribute in the data. We assign the record pair \((i,j)\) to a
\emph{linkage cluster} \(S_k \in \{S_1, \ldots, S_K\}\) only when both
records belong the corresponding record clusters; otherwise we assign it
to \(S_0\). For example
\[i \in A_1, j \in B_1 \implies \gamma_{ij} \in S_1\]
\[i \in A_1, j \in B_2 \implies \gamma_{ij} \in S_0\] Cluster
assignments at the record pair level are necessary for the updating the
\(\mathbf{m}\) and \(\mathbf{u}\) parameters. We denote the reliability
and discernment parameters assoicated with each field for each cluster
as \(\mathbf{m}^{fk}\) and \(\mathbf{u}^{fk}\) respectively. To improve
readability, we modify the definition of \(\mathbf{m}_{ij}\) to
incorporate the linkage cluster assignment. Explicitly,
\[\mathbf{m}_{ij} = \prod_{f=1}^{F}\prod_{l=1}^{L_f}\prod_{k=0}^{K}\left[ \left(m^{flk}\right)^{\mathbf{1}_{\gamma_{ij} \in S_k}}\right]^{\mathbf{1}_{\gamma_{ij}^f = l}}\]
with an analogous definition for \(\mathbf{u}_{ij}\).

Additionally, define the vector of linkage probabilities
\(\Lambda_k = (\lambda_{k1}, \lambda_{k2}, \lambda_{k3})\), where
\(\lambda_{k1} = P(z_{j\in S_k} \in S_k)\) is the probability that a
record in \(B_k\) has a match within its cluster in \(A_k\),
\(\lambda_{k2} = P(z_{j \in S_k} \in S_{k'})\) is the probability that
such a record \(B_k\) has a match outside of its cluster, in
\(\mathbf{X}_{A_{k'}}\) for some \(k' \neq k\), and
\(\lambda_{k3} = P(z_{j\in S_k} = n_A + 1)\) is the probability of a
non-match. This is a multinomial distribution where we necessarily have
\(\lambda_{k3} = 1 - \lambda_{k1} - \lambda_{k2}\). These parameters use
linkage cluster assignment at the record level.

Our new model specification becomes

\[\mathbf{m^{fk}} \sim \text{Dirichlet}(\alpha^{f1k}, \ldots, \alpha^{fL_fk})\]

\[\mathbf{u^{fk}} \sim \text{Dirichlet}(\beta^{f1k}, \ldots, \beta^{fL_fk})\]
\[Z_{j \in S_k} | \Lambda_k =
\begin{cases} 
    \frac{1}{n_{A_k}}\lambda_{k1}  & z_j \in S_k; \\
    \frac{1}{n_A - n_{A_k}}\lambda_{k2}  & z_j \in S_{k'}; \\
     \lambda_{k3} & z_j = n_A + 1 \\
\end{cases}\]

\[\Lambda \sim \text{Dirichlet}(\Lambda_0)\] where
\(\Lambda_0 = (\Lambda_{k1}, \Lambda_{k2}, \Lambda_{k3}).\)

\textbf{THIS WOULD GO IN APPENDIX}: Through similar derivations, this
yields full conditionals: The \(\mathbf{m}\) and \(\mathbf{u}\)
parameters are updated through standard multinomial-dirichlet mechanics.
Thus we have
\[\mathbf{m^{fk}}|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\alpha^{f1k}(\mathbf{Z}), \ldots, \alpha^{fL_fk}(\mathbf{Z}))\]

\[\mathbf{u^{fk}}|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\beta^{f1k}(\mathbf{Z}), \ldots, \beta^{fL_fk}(\mathbf{Z}))\]
where
\[\alpha_{flk}(\mathbf{Z})= \sum_{i,j, k} \mathbf{1}_{obs(\gamma_{ij}^f)}\mathbf{1}_{\gamma_{ij}^f = l} \mathbf{1}_{z_j = i}\mathbf{1}_{\gamma_{ij} \in S_k}\]

\[\text{and } \beta_{flk}(\mathbf{Z})= \sum_{i,j, k} \mathbf{1}_{obs(\gamma_{ij}^f)}\mathbf{1}_{\gamma_{ij}^f = l} \mathbf{1}_{z_j \neq i}\mathbf{1}_{\gamma_{ij} \in S_k}\]

We also have straightforward posterior distritutions for the linkage
probability vector \(\Lambda_k\):

\[\Lambda_k | \mathbf{Z} \sim \text{Dirichlet}(\mathbf{D_k} + \Lambda_0)\]

where \(\mathbf{D}_k = (D_{k1}, D_{k2}, D_{k3})\) with
\[D_{k1} = \sum_{j \in B_k} \mathbf{1}_{z_j \in A_k}\]
\[D_{k2} = \sum_{j \in B_k} \mathbf{1}_{z_j \in A_{k'}}\]
\[\text{and } D_{k3} = \sum_{j \in B_k} \mathbf{1}_{z_j = n_A + 1}\]

Our full conditional for \(Z_j\) follows a familiar form:

\[Z_{j \in S_k}^{(s+1)} | \mathbf{m}^k, \mathbf{u}^k, \Gamma, \Lambda_k =
\begin{cases} 
    \frac{\lambda_{k1}}{n_{A_k}} w_{ij} & z_j \in S_k; \\
    \frac{\lambda_{k2}}{n_A - n_{A_k}} w_{ij}  & z_j \in S_{k'}; \\
     \lambda_{k3} & z_j = n_A + 1 \\
\end{cases}\]

And after integrating over \(\Lambda_k\) and rearranging terms,

\[Z_{j \in S_k}^{(s+1)} | \mathbf{m}^k, \mathbf{u}^k, \Gamma,  =
\begin{cases} 
    \frac{D_{k1} + \Lambda_{01}}{n_{A_k}(n_B - D_{k1} - D_{k2} + \Lambda_{03})} w_{ij} & z_j \in S_k; \\
    \frac{D_{k2} + \Lambda_{02}}{(n_A - n_{A_k})(n_B - D_{k1} - D_{k2} + \Lambda_{03})} w_{ij}  & z_j \in S_{k'}; \\
     1 & z_j = n_A + 1 \\
\end{cases}\]

>>>>

\setstretch{1}
\begin{flushright} 
    \end{flushright}
    \begin{center} \textbf{Parlr: Parallelized Record Linkage in R}

@@

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

Most methods for conducting record linkage, the task of identifying
duplicate records across datasets, are derived from the foundational
article by Fellegi and Sunter in 1969. Within this family of methods,
two methods have notable strengths: \texttt{fastlink}, proposed by
Enamorado et al (2019), continues a is able to conduct linkage on
incredibly large datasets

@@

(This will be revised to talk more explicitly about sampling \(Z_j\)
independently)

@@

\[\lambda \sim \text{Beta}(\alpha_{\lambda}, \beta_{\lambda}) \] The
prior for \(Z_j\) has equal probability of matching to all records
\(i\in A\), and non-matching probability governed by \(\lambda\).
Therefore a \(\lambda \sim \text{Beta}(1, 1)\) corresponds to a prior
belief that nonmatches and matches are equally likely, and a
\(\lambda \sim \text{Beta}(1, \frac{1}{n_A})\) prior corresponds to a
uniform prior on the labeling of \(\mathbf{Z}\).

@@

\underline{Sample $\lambda^{(s+1)}|\mathbf{Z}^{(s)}$:} As a function of
\(\lambda\), the linkage structure parameter \(\mathbf{Z}\) is sequence
of successes (when \(z_j < n_A + 1\)) and failures (when
\(z_j = n_A + 1\)), and therefore
\(p(\mathbf{Z}|\lambda) = \mathcal{L}(\lambda|\mathbf{Z})\) is
determined only by the number of duplicates
\(D = \sum_{i=1}^{n_B}\mathbf{1}_{z_j < n_A + 1}\) encoded by
\(\mathbf{Z}\). Thus we have

@@

\underline{Sample $\mathbf{Z}^{(s+1)}|\Gamma, \mathbf{m}^{(s+1)}, \mathbf{u}^{(s+1)}, \lambda^{(s+1)}$:}
Because we sample \(Z_j\) independently of all other \(Z_{j'}\), we use
only the full conditional for an individual \(Z_j\). Let \(\Gamma_{.j}\)
denote the set of \(n_A\) comparison vectors with \(j \in B\), and note
that as a function of \(Z_j\), the likelihood
\(p(\Gamma_{.j}|Z_j, \mathbf{m}, \mathbf{u}) = \mathcal{L}(Z_j|\Gamma_{.j}, \mathbf{m}, \mathbf{u})\)
is a discrete distribution with probabilities proportional to

@@

In order to make fair comparisons against the Sadinle 2017 model, we
integrate over the posterior of \(\lambda\) and rearrange terms to
produce the final full conditional:

\[p\left(Z_j^{(s+1)}  = i| \mathbf{m}, \mathbf{u}, \mathbf{Z^{(s)}}\right) \propto
\begin{cases} 
    w_{ij}  & i \leq n_A; \\
     n_A \frac{n_B - D + \beta_{\lambda}}{D + \alpha_{\lambda}} & i  = n_A + 1 \\
\end{cases}\]

@@

Our Gibbs sampler provides posterior samples of \(\mathbf{Z}\) which we
use to make our final decisions about the linkage structure. In the case
where we false matches and missed matches contribute the same loss, we
declare \((i,j)\) to be a match whenever \(P(Z_j = i) > \frac{1}{2}\)
according to these posterior samples. We can create more elaborate
decisions by attributing different loss values to different kinds of
error, and allowing pairings with middling posterior probabilities to be
left unlabeled by the algorithm so that the modeler can address those
manually. For further discussion, see Sadinle 2017.

\hypertarget{efficient-computation}{%
\section{Efficient Computation}\label{efficient-computation}}

Broadly speaking, we increase our computational efficiency by
recognizing that record pairs contribute to posterior calculations only
through the agreement pattern of the \(\gamma_{ij}\) vector. Let \(H\)
be the set of unique agreement patterns in the data, let \(P\) denote
the total number of unique agreement patterns. Note that \(P\) is
bounded above by \(\prod_{f=1}^F L_f\), and that this bound does not
scale with \(n_A\) or \(n_B\). We index these agreement patterns by
\(p \in \{1, \ldots, P\}\), and say \((i,j) \in h_p\) when the \((i,j)\)
pair exhibits the \(p^{th}\) agreement pattern. Wherever possible, we
conduct calculations over these \(P\) agreement patterns rather than the
\(n_A \times n_B\) record pairs.

\hypertarget{data-representation-and-hashing}{%
\subsection{Data Representation and
Hashing}\label{data-representation-and-hashing}}

In the classic Fellegi Sunter framework, \(\Gamma\) is a
\(n_A n_B \times F\) matrix, with each row providing the comparison
vector for a different \((i,j)\) pair. We however do not store these
comparison vectors themselves, but instead only the a hashed value
corresponding to the agreement pattern of the \((i, j)\) pair. Enamorado
et al (2019) provided the hashing function
\[\sum_{f=1}^F \mathbf{1}_{\gamma_{(i,j)}^f >0}2^{\gamma_{(i,j)}^f + \mathbf{1}_{f>1} \times \sum_{e=1}^{f-1}(L_e-1)}\]
to map each agreement pattern to a unique value, but packages like
\texttt{dplyr} in \texttt{R} are capable of this as well.

We store this information in a nested list \(\tilde{\Gamma}\) where the
\(p^{th}\) component of the \(j^{th}\) list contains a vector of records
in \(A\) that share agreement pattern \(h_p\) with record \(j \in B\).
For each \(p\), we also calculate
\(|h_p| = \sum_{i=1}^{n_A}\sum_{j=1}^{n_B} \mathbf{1}_{(i,j) \in h_p}\)
total instances of agreement pattern \(h_p\) throughout the data, and
also for each \(j\), we calculate
\(|h_p|_j = \sum_{i=1}^{n_A} \mathbf{1}_{(i,j) \in h_p}\) the instances
of agreement pattern \(p\) among the comparison vectors between record
\(j \in B\) and each of the \(n_A\) records in \(A\).

For large data, we can partition the two datasets \(A\) and \(B\) into
smaller blocks \(\{A_m\}\) and \$\{B\_m\} for more manageable
computations. On a single machine, we can read-in data sequentially,
conduct hashing, collect results, and delete the original data from
memory before continuing with the next chunk of data. With multiple
cores or multiple machines, this can be done in parallel. Storing this
hashed information still becomes burdensome for large data, but this
hashing method greatly expands the capabilities of the Fellegi-Sunter
framework.

Lastly, the classic Fellegi Sunter method represents the \(\gamma_{ij}\)
comparison vector as vector of length \(F\), with each component
\(\gamma_{ij}^f\) taking on values in \(\{0, \ldots, L_f - 1\}\). To
ease computations, we instead represent the comparison as a
concatenation of \(F\) many binary indicator vectors of lengths \(L_f\).
For example, if \(L_1 = L_2 = 2\) and \(L_3 = 3\), then
\(\gamma_{ij} = (1, 0, 2)\) under the classical framework becomes
\(\gamma_{ij} = (0, 1, 1, 0, 0, 0, 1)\) under our framework. This is a
bijective transformation that does not change the meaning of the data,
but this representation will ease calculations and posterior updates.

\hypertarget{gibbs-sampling}{%
\subsection{Gibbs Sampling}\label{gibbs-sampling}}

After receiving matching statuses from \(\mathbf{Z}\), the Sadinle
method calculates \(\alpha_{fl}(\mathbf{Z})\) and
\(\beta_{fl}(\mathbf{Z})\) for each field and level. This constitutes
\(2 \times \sum L_f\) many summations over \(n_A \times n_B\)
quantities, and becomes computationally burdensome with large data. In
contrast, we recognize that each unique agreement pattern contributes to
the posterior \(\alpha(\mathbf{Z})\) and \(\beta(\mathbf{Z})\) vectors
in the same way. In fact, if we denote
\(\mathbf{Z}(h_p) = \sum_{j=1}^{n_B} \mathbf{1}_{(Z_j, j) \in h_p}\) to
be the number of matching record pairs with agreement pattern \(h_p\),
then the contribution of pairs of pattern \(h_p\) to the
\(\alpha(\mathbf{Z})\) vector is simply \(\mathbf{Z}(h_p) \times h_p\).
Thus our posterior updates are simply
\(\alpha(\mathbf{Z}) = \alpha_0 + \sum_{p=1}^P \mathbf{Z}(h_p) \times h_p\)
and
\(\beta(\mathbf{Z}) = \beta_0 + \sum_{p=1}^P |h_p| - \left(\mathbf{Z}(h_p) \times h_p \right)\).
Note that this constitutes \(P\) many summations over \(n_B\)
quantities, and thus avoids the \(n_A \times n_B\) summation from the
original method.

Sadinle uses a prior for \(\mathbf{Z}\) that induces the a full
conditional for \(Z_j\) that strictly enforces one-to-one matching.
Particular, this sampler removed previously matches records from the set
of candidate records when sampling \(Z_j\), creating a dependency that
makes the sampler \emph{inherently serial}.By weakening the one-to-one
requirement, our full conditional for \(Z\) does not depend on the rest
of the \(\mathbf{Z}_{-j}\) vector, and thus can be computed in parallel.
More importantly, since only the agreement pattern of \(Z_j\) is used
for calculations within the Gibbs sampler, and not the particular record
label, we can conduct this sampling only at the level of the unique
agreement patterns. This boosts computation time far greater than
parallelization.

To do this we calculate the Fellegi Sunter weight \(w_{h_p}\) for each
unique pattern and sample the agreement pattern \(h(Z_j)\) following the
modified full conditional:

\[h\left(Z_j^{(s+1)}\right) | \mathbf{m}, \mathbf{u}, \mathbf{Z^{(s)}} \propto
\begin{cases} 
    w_{h_p}\times |h_p|_j  & h(z_j) \in H; \\
     n_A \frac{n_B - D + \beta_{\lambda}}{D + \alpha_{\lambda}} &  z_j  = n_A + 1 \\
\end{cases}\]

We complete the entire Gibbs procedure at the level of the \(P\)
agreement patterns. After, we can back-fill the records corresponding to
the agreement patterns by sampling uniformly at random among candidate
records stored in \(\tilde{\Gamma}\).

\hypertarget{enforcing-bipartite-matching}{%
\subsection{Enforcing Bipartite
Matching}\label{enforcing-bipartite-matching}}

Add this

\hypertarget{simulation-studies}{%
\section{Simulation Studies}\label{simulation-studies}}

\begin{itemize}
\item
  Replicating the Sadinle Study
\item
  \texttt{parlr} and \texttt{BRL} speed comparison
\end{itemize}

\hypertarget{data-analysis}{%
\section{Data Analysis}\label{data-analysis}}

\begin{itemize}
\tightlist
\item
  Necessary? El Salvador data? Or perhaps needs to be a bigger dataset?
\end{itemize}
